ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-07 rebuttal.tex

\usepackage[usenames, dvipsnames]{xcolor}
\usepackage{siunitx}
\sisetup{group-separator={,},
     detect-all,
     binary-units,
     list-units = single,
     range-units = single,
     range-phrase = --,
     per-mode = symbol-or-fraction,
     separate-uncertainty = true,
     multi-part-units = single,
     list-final-separator = {, and }
%    scientific-notation = fixed
}
\providecommand{\hly}[1]{\colorbox{yellow}{#1}}
\providecommand{\hlb}[1]{\colorbox{SkyBlue}{#1}}
\providecommand{\hlg}[1]{\colorbox{green}{#1}}

% bump up section font size
\usepackage{sectsty}
\sectionfont{\fontsize{12}{15}\selectfont}

\usepackage[colorlinks=true]{hyperref}
\usepackage{todonotes}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.25cm}

\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup}
\newenvironment{response}{\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
\newcommand{\review}[1]{{\itshape#1}}
\newcommand{\respond}[1]{\textbf{Our Response:} #1}

@@

\begin{reviewer}
    %
    p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind
    considerably, although use of the former continually grows due to Cantera's open availability.''

    Note that FlameMaster is also freely available from the GitLab repository of the developers.
    %
\end{reviewer}

\begin{response}
    We attempted to access the GitLab repository at \url{https://git.rwth-aachen.de/ITV/FlameMaster}, but
    it is a closed repository. Thus, while it may be freely available to those with access, it is not
    \emph{openly} available.

    To clarify the meaning of ``open'' when used in our article, we have added to \textbf{page 2} a
    brief definition and link to the definition of open source, which comes from the
    Open Source Initiative.
\end{response}

@@

    We have modified the text on \textbf{page 2} to clarify this point, including describing the lack of human
    readability as a formatting issue (but one that is inherent to XML), rather than a technical
    issue.
    %
\end{response}

\begin{reviewer}
    %
    I suggest adding to following block to this para:

    ``The original PrIMe was on an Amazon cloud server. Recently it was shifted to the computer
    center of DLR Stuttgart. At the new location it is called ``PrIMe 2.0''
    [\url{https://prime.cki-know.org/}]. This is identical to the original PrIMe apart from
    technical changes due to the different computation background. Professor Frenklach plans to
    develop further PrIMe (using name PrIMe 3.0) and this version will be availabe at
    [\url{http://primekinetics.org/}].''
    %
\end{reviewer}

\begin{response}
    %
    We have added the new PrIMe URL to the manuscript on \textbf{page 2}, and thank the reviewer for pointing us to that.

    However, we do not think that the technical details of the implementation of the original PrIMe system
    are relevant to this work, so we have not included them. In addition, we cannot know
    the plans of Prof.~Frenklach, and we do not feel it would be appropriate to comment on them in
    our manuscript. As such, we have not added these suggestions to the paper.
    %
\end{response}

@@

\begin{reviewer}
    %
    5.~Page 1, line 31: The future development plans include support for jet stirred reactors,
    however, this is not mentioned in section 5, Conclusions and future work, nor is there any
    reference to it on the GitHub wiki roadmap of PyKED.
    %
\end{reviewer}

\begin{response}
    %
    Thank you for the comment, we have corrected this oversight in the roadmap and the conclusions.
    %
\end{response}

\begin{reviewer}
    %
    6. I don't completely agree with the second part of the statement on page 2 line 16-18:
    ``Second, the XML format is intended to be a machine-readable mark-up language rather than a
    data format, and its lack of human readability presents a barrier to creating and working with
    database files.''

    In my opinioo n the YAML format resembles the XML format strongly. XML format is human-readable,
    however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to
    read. If XML would lack human readability as it is stated in the article, it would not be such a
    popular format, and I don't think PrIMe and ReSpecTh would be using it.
    %
\end{reviewer}

### 2017-10-07 rebuttal.tex

\begin{reviewer}
    %
    ``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for
    typical bibliographic data.''

    I suggest the following extension:

    ``For example, where PrIMe uses internal bibliographic references, ReSpecTh adds a field for
    typical bibliographic data, including the DOI of the cited arecicle. Each datafile has a unique
    own DOI and unique URL address.
    [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}]''. Also species can be
    identified by InCHI or SMILES.
    %
\end{reviewer}

\begin{response}
    %
    We have incorporated these modifications into the suggested paragraph on \textbf{page 2},
    although we have modified the format slightly to better fit the flow of the text.
    %
\end{response}

### 2017-10-07 rebuttal.tex

\begin{reviewer}
    %
    ``The ReSpecTh standard also provides machine-readable formats for describing ignition
    experiments, including a field for the definition of autoignition and the ability to specify
    facility-specific effects.''

    Modified text:

    ``The ReSpecTh standard provides formats that are both human and machine-readable for describing
    ignition experiments, including a field for the definition of autoignition and the ability to
    specify facility-specific effects. Also, it describes PSR and flow reactor speciation data,
    burner stabilized flame speciation data and burning velocity data.''
    %
\end{reviewer}

\begin{response}
    %
    We have incorporated most of the recommended text in our revision on \textbf{page 2}.
    However, we did not add ``human-readable'', consistent with the definition used throughout
    and explained on \textbf{page 2} (i.e., the XML file is more verbose and less easily understood).
    %
\end{response}

### 2017-10-07 rebuttal.tex

\begin{response}
    %
    We have incorporated most of the recommended text in our revision on \textbf{page 2}.
    However, we did not add ``human-readable'', consistent with the definition used throughout
    and explained on \textbf{page 2} (i.e., the XML file is more verbose and less easily understood).
    %
\end{response}

\begin{reviewer}
    %
    ``However, ReSpecTh experimental files are another XML-based format, and, as such, suffer from
    the same usability issues as PrIMe.''

    I think that the limits of an XML database is set by the actual set of tags and not the format.
    %
\end{reviewer}

\begin{response}
    %
    By ``usability'' of a format, we mean whether files are easily usable and understandable by
    people, rather than the technical limitations of a format. We have changed the word
    ``usability'' to ``readability'' on \textbf{page 2} to emphasize this.
    %
\end{response}

@@

\begin{reviewer}
    %
    ``Moreover, the standard remains closed, requiring potential users to register with the
    standard's authors to access the specification and the growing database.''

    I suggest instead:

    ``Registration is open for everyone who provides a valid institutional e-mail address.''
    %
\end{reviewer}

\begin{response}
    %
    We have updated the text on \textbf{page 2} to reflect this information; however, we argue
    that requiring registration to \emph{view} the specification and data limits use
    of the standard and differs from the openness that we are advocating (now defined on \textbf{page 2}),
    where the specification and data files are open for anyone to view, with or without
    registration.
    %
\end{response}

### 2017-10-07 rebuttal.tex

\begin{reviewer}
    %
    ``... hydrogen/syngas [11] and ethanol [12]. In all four cases, they converted numerous
    experimental datasets from the literature into the ReSpecTh standard.''

    Extended version:

    ``... hydrogen/syngas [11], methanol and formaldehyde [C. Olm, T. Varga, E. Valko, H. J. Curran,
    T. Turanyi: Uncertainty quantification of a newly optimized methanol and formaldehyde combustion
    mechanism, Combustion and Flame, in press] and ethanol [12]. In all five cases, they converted
    numerous experimental datasets from the literature into the ReSpecTh standard. The database
    currently contains  93240 experimental data points in 1376 XML-format data files
    [\url{http://respecth.chem.elte.hu/respecth/reac/CombustionData.php}].''
    %
\end{reviewer}

\begin{response}
    We added the suggested text and reference to \textbf{page 3}.
\end{response}

@@

\begin{reviewer}
    %
    ``type (required, string): How ignition delay was measured;''

    The options described here is a very limited subset of the ignition delay time definitions used
    in experimental articles. For example, backward or forward interpolated values of a measured
    quantity are further possibilities.
    %
\end{reviewer}

\begin{response}
    %
    We have implemented an option for baseline extrapolation of the slope, and added the
    description of this ignition type to \textbf{page 6} of the manuscript. To our knowledge,
    the options now supported represent all of the common methods of specifying the ignition delay
    in rapid compression machine and shock tube experiments.
    %
\end{response}

### 2017-10-07 rebuttal.tex

\begin{response}
    %
    We have implemented an option for baseline extrapolation of the slope, and added the
    description of this ignition type to \textbf{page 6} of the manuscript. To our knowledge,
    the options now supported represent all of the common methods of specifying the ignition delay
    in rapid compression machine and shock tube experiments.
    %
\end{response}

\begin{reviewer}
    %
    p.~8 ``In the latter case, some information may be lost, as ChemKED files both encode more
    details and specify more granularity of some information than ReSpecTh files.''

    This was true for ReSpecTh format 1.0, but no longer true for ReSpecTh Kinetic Dataformat 2.0.
    The latter covers all info of the ChemKED files (and more).
    %
\end{reviewer}

\begin{response}
    %
    This is still the case, even in ReSpecTh format 2.0. For instance, ChemKED files support
    specifying the initial composition as a mass fraction, which ReSpecTh does not support. We have
    modified the sentence on \textbf{page 10} to indicate that there are some incompatibilities
    in both directions, not all ChemKED files can be converted to ReSpecTh files, and
    not all ReSpecTh files can be converted to ChemKED files due to missing data elements in both directions.
    %
\end{response}

### 2017-10-07 pyked.tex
minor edits for #15

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/}
has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML can represent detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be much more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them. Furthermore, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

@@

The \pybox|write_file()| method writes a new \ck{} YAML file based on the instance attributes
of the class, while \pybox|convert_to_ReSpecTh()| converts the \pybox|ChemKED|
instance to a ReSpecTh XML file. In the latter case, some information may be lost or the conversion
may not be possible, as \ck{} files support some data that ReSpecTh files do not support. We note
that the reverse is also true, not all ReSpecTh files may be converted to \ck{} files because not
all of the data that can be stored in a ReSpecTh format is supported by the \ck{} format.

>>>>

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/})
has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML can represent detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be much more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them. Furthermore, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

@@

The \pybox|write_file()| method writes a new \ck{} YAML file based on the instance attributes
of the class, while \pybox|convert_to_ReSpecTh()| converts the \pybox|ChemKED|
instance to a ReSpecTh XML file. In the latter case, some information may be lost or the conversion
may not be possible, as \ck{} files support some data that ReSpecTh files do not support. We note
that the reverse is also true: not all ReSpecTh files may be converted to \ck{} files because not
all of the data that can be stored in a ReSpecTh format is (currently) supported by the \ck{} format.
